Define the tangent space of a manifold

The tangent space is a linear (more rarely, affine) space of vectors tangent to a smooth (differentiable) manifold \( M \) at a given point \( a \in M \),
denoted by \( T_a M \)

For curves, surfaces and submanifolds embedded in a Euclidean subspace \( \mathbb{R}^n \) the tangent subspace can be identified with an affine subset 
(of the corresponding dimension \( 1,2,\dots \)) in the ambient space, passing through \( a \). 
For abstract manifolds \( T_aM \) is naturally isomorphic to the linear space of derivations \( D : C^\infty(M) \to \mathbb{R}^m \) of 

of the ring of smooth functions on \( M \) satisfying the Leibniz rule:
\(  D\in T_a M\iff D:C^\infty(M)\to\R\) 
\(D(f\pm g)=Df\pm Dg,\ D(\lambda f)=\lambda Df\) 
\(D(f\cdot g)=f(a)\cdot Dg+g(a)\cdot Df. \)

The main reason for the abstraction is that we do not have any ambient space in which we could define the vectors tangent to the manifold.
The (disjoint) union of tangent spaces to a manifold to all its points has a natural structure of a over \( M \), called the tangent bundle.

To make the relation between the affine space of vectors tangent to a point and the ring of derivations more concrete,
say \( M \subset \mathbb{R}^m \) and take any \( v \in \mathbb{R}_{a}^m \) (vectors tangent to the point \( a \)) 
and \( f \in C^{\infty}(\mathbb{R}^n) \) then we can build the directional derivative at \( a \) along \( v \) via

\( D_v(f)|_a = D_v f(a) = \lim_{h \to 0} \frac{f(a + hv) - f(a)}{h} = \frac{d}{dt} f(a + tv)|_{t=0} \)
this is linear and satisfies of course the Lebniz rule.
If we write \( v_a = v^i e_i|_a \) in a standard basis the chain rule gives
\( \frac{d}{dt} f(a + tv)|_{t=0} = v(\frac{df}{dx}(a + tv))|_{t=0} = v^i\frac{df}{dx^i}(a) \).

In fact one can prove that there is an isomorphism of \( v \in T_a M \) and derivations \( D_v \) at \( a \),
but the latter exists without any prior assumption of \( M \) being in \( \mathbb{R}^m \).

Conversely one can interpret \( D(f) : \mathbb{R}^m \to \mathbb{R}^n \) (as was it done in the lecture), although this requires
additional effort because \( \mathbb{R}^m \) actually refers here to a local parameterization of \( f \) via some chart
\( (\phi, U) \) and actually \( f : U \to \mathbb{R}^n \) such that \( D(f) \coloneqq D(\tilde{f} \) and \( \tilde{f} = f \circ \phi^{-1} \).

This causes then the issue of verifying compatibility with transition maps, that is our tangent vectors \( v \) are taken from the tangent space 
to \( \phi(a) \in \mathbb{R}^m \) and one has to show that if \( (\psi, V) \), \( a \in V \) and any \( f : U \to \mathbb{R}^n \)
parameterized via \( \phi \) by default and \( \tilde{f} \) its parameterization via \( \psi \):

\( Df(v)|_a = D(f \circ \phi^{-1})(v)|_a = D(f \circ \psi^{-1} \circ \psi \circ \phi^{-1})(v)|_a = D(\tilde{f} \circ \psi \circ \phi^{-1})(v)|_a \)
\( = D\tilde{f})(D \psi \circ \phi^{-1}(v))|_{\psi(a)}\)

so if \( Df(v) = D\tilde{f}(\tilde{v})\) then it must be \( \tilde{v} = D \psi \circ \phi^{-1}(v) \).
Thus the tangent space at \( a \) is here defined as the quotient space of all \( v \in T_{\phi(a)}, \tilde{v} \in T_{\psi} \)
s.t.   
\( \tilde{v} = D(\psi \circ \phi^{-1})(v)|_{a} \)


Define the differential of a map \( F : M \to N \) at \( p \)

Assuming that \( F \) is \( C^k \) differentiable at \( p \) the differential can be defined via the derivation construction
as the map \( dF_p : T_p M \to T_{F(p)} N\) that is  \( dF_p \) refers to the derivation in \( T_p N \) satisfying
for \( f \in C^k(N)\) and derivation \( (v : C^k(M) \to \mathbb{R}) \in T_p M \)
\( dF_p(v)(f) = v(f \circ F) \)
since \( f \circ F \in C^k(M) \), so \( v(f \circ F) \) is well defined.

Properties of Differentials. 
Let \(M\), \(N\), and \(P\) be smooth manifolds with or without boundary, let \(F: M \to N\) and \(G: N \to P\) be smooth maps, and let \(p \in M\).
(a) \(dF_p: T_pM \to T_{F(p)}N\) is linear.
(b) \(d(G \circ F)_p = dG_{F(p)} \circ dF_p: T_pM \to T_{G \circ F(p)}P\).
(c) \(d(\text{Id}_M)_p = \text{Id}_{T_pM}: T_pM \to T_pM\).
(d) If \(F\) is a diffeomorphism, then \(dF_p: T_pM \to T_{F(p)}N\) is an isomorphism, and \((dF_p)^{-1} = d(F^{-1})_{F(p)}\).

With that its also relatively easy to obtain the lecture definition (which is again dependent on local charts).
For any chart \( \phi : U \to \mathbb{R}^m \) we have that \( \{\frac{d}{dx^i}|_{\phi(p)}\}_{i \in [m]} \) forms a basis of \( T_{\phi(p)} \mathbb{R}^m \)
and as \( \phi \) is a diffeomorphism \( d\phi_p : T_p M \to T_{x(p)} \mathbb{R}^m \) is a isomorphism between and for any \( f \in C^k(U) \)
we can define the action of a basis vector \( \frac{d}{d\tilde{x}^i}|_p = (d\phi_p)^{-1}(\frac{d}{dx^i}|_{\phi(p)}\) of \( T_p M \) via
\( \frac{d}{d\tilde{x}^i}|_p f = \frac{d}{dx^i}|_{\phi(p)}(f \circ \phi^{-1}) \).
More generally if \( v = v_i \frac{d}{d\tilde{x}^i} \in T_p M \) and \( F : M \to N \) then
\( dF_p(v) = dF_p(v_i \frac{d}{d\tilde{x}^i}) = v_i dF(\frac{d}{d\tilde{x}^i}) = v_i \frac{dF^j}{d\tilde{x}^i}\frac{d}{d\tilde{y^j}} \)
Note that we are still using the Einstein sum notation so 
\( v_i \frac{dF^j}{d\tilde{x}^i}\frac{d}{d\tilde{y^j}} = \sum_i^m \sum_j^n v_i \frac{dF^j}{d\tilde{x}^i}\frac{d}{d\tilde{y^j}} \)
since \(dF_j \) generally does not map one basis vector \( \frac{d}{d\tilde{x}^i} \) to some \( \frac{d}{d\tilde{y}^i} \) but rather
a linear combination of the \( \frac{d}{d\tilde{y}^i} \). In particular we can see that this is nothing but the Jacobi matrix in disguise
with each column encoding how \( \frac{d}{d\tilde{x}^i} \) translates into \( \frac{d}{d\tilde{y}^i} \).

State how the basis vectors of the tangent space behave under a change of coordinates

As we have seen for the description of the differential, if \( \{\frac{d}{d\tilde{x}^i}\} \) is the basis
induced by the chart \( (U, \phi) \) via \( \frac{d}{d\tilde{x}^i} = d\phi^{-1}(\frac{d}{d\tilde{x}^i}) \in T_{\phi(p)} \mathbb{R}^k \) 
at the point \( p \) and \( \{\frac{d}{d\tilde{y}^i}\} \) for a chart \( (V, \psi) \) and for ease of notation let \( F = \psi \circ \phi^{-1} \)
then by the Jacobian representation of the differential

\(d_{\phi(p)}(\psi \circ \phi^{-1})(\frac{d}{dx^i}|_{\phi(p)} = \frac{dF^j}{dx^i}(\phi(p))\frac{d}{dy^j} \)

put even more succinct let be two local representations of the same tangent vector be \( v = v^i\frac{d}{dx^i} = = w^i\frac{d}{dy^i} \) 
(living in \(\mathbb{R}^m\)) 
we have that 
\( w^j = \frac{dF^j}{dx^i}(\phi(p))v^i \)
which is simply the sum of the entries in the \(j\)-th row of the Jacobian.

\( d_{\phi(p)}(\psi \circ \phi^{-1})(\frac{d}{dx^i}|_{\phi_p} = \frac{d\tilde{x}^j}{dx^i}(\phi(p))\frac{d}{d} \)

A vector bundle \(E\) over a base \(M\). 
A point \(m_1\) in \(M\) corresponds to the origin in a fibre \(E_{m_1}\) of the vector bundle \(E\), and this fibre is mapped down to the point \(m_1\) by the 
projection \(\pi: E \to M\).]]

Define a vector bundle

A real vector bundle consists of:
1. topological spaces \(X\) base space and \(E\) total space.
2. a continuous surjection \(\pi:E \to X\) bundle projection.
3. for every \(x\) in \(X\), the structure of a finite-dimensional real vector space on the fiber \(\pi^{-1}(\{x\}\))

where the following compatibility condition is satisfied: 
for every point \(p\) in \(X\), there is an open neighborhood \(U\subseteq X\) of \(p\), a natural number \(k\), and a homeomorphism
\(\varphi\colon \pi^{-1}(U) \to U \times \mathbb{R}^k\) 


such that for all \(x\) in \(U\) and \( v \in \pi^{-1}(x) \),
* \( \text{pr}_1 \circ \varphi(v) = \pi(v) = x \) for all vectors \(v\) in \(\mathbb{R}^k\)
* The restriction of \( \phi \) to \( \pi^{-1}(x) \) is a vector space isomorphism from \( \pi^{-1}(x) \) to \( \{q\}\times \mathbb{R}^k \simeq \mathbb{R}^k \)

Pictorially the following diagram commutes

We refer to \( (U, \phi) \) as  a local trivialization and intuitively says the locally the space \( E \) looks like \(U\times \mathbb{R}^k\).
If there exists a set \( \{(U_i, \phi_i)\} \) with such homeomorphism and \( U_i \) covering all of \( E \), the vector bundle is called trivial.

A generalization are fiber bundles, where we have a triple \( (E, X, \pi, F) \) where \( E \) is again a total space,
\( X \) the base space, but this time we do not require \( F = \mathbb{R}^k \) but any topological space, 
so we do not have necessarily a vector space structure locally.


Describe the special case of the Tangent bundle for vector bundles
and show that it is a manifold

Let the total space \( E = TM = \amalg_{p \in M} T_p\ M \) and the natural projection map \( \pi : TM \to M,\ \pi(v_p) = p \)
where \( v_p \) is a tangent vector at \( p \). The local trivialization is given by \( \phi : \pi^{-1}(U) \to U \times \mathbb{R}^m \)
via \( \phi(v_p) = (p, v) \).

For it to be a manifold, we begin by defining the maps that will become our smooth charts. 
Given any smooth chart \((U, \varphi)\) for \(M\), note that \(\pi^{-1}(U) \subset TM\) 
is the set of all tangent vectors to \(M\) at all points of \(U\). 
Let \((x^1, \dots, x^n)\) denote the coordinate functions of \(\varphi\), 
and define a map \(\tilde{\varphi} : \pi^{-1}(U) \to \mathbb{R}^{2n}\) by 
\(\tilde{\varphi} \left( \left. v^i \frac{\partial}{\partial x^i} \right|_p \right) = (x^1(p), \dots, x^n(p), v^1, \dots, v^n). \qquad (3.13)\)

Its image set is \(\varphi(U) \times \mathbb{R}^n\), which is an open subset of \(\mathbb{R}^{2n}\). 
It is a bijection onto its image, because its inverse can be written explicitly as
\(\tilde{\varphi}^{-1}(x^1, \dots, x^n, v^1, \dots, v^n) = \left. v^i \frac{\partial}{\partial x^i} \right|_{\varphi^{-1}(x)}.\)

Now suppose we are given two smooth charts \((U, \varphi)\) and \((V, \psi)\) for \(M\), 
and let \((\pi^{-1}(U), \tilde{\varphi}), (\pi^{-1}(V), \tilde{\psi})\) be the corresponding charts on TM. 

The sets 
\(\tilde{\varphi}(\pi^{-1}(U) \cap \pi^{-1}(V)) = \varphi(U \cap V) \times \mathbb{R}^n\) and
\(\tilde{\psi}(\pi^{-1}(U) \cap \pi^{-1}(V)) = \psi(U \cap V) \times \mathbb{R}^n\)

are open in \(\mathbb{R}^{2n}\), and the transition map 
\(\tilde{\psi} \circ \tilde{\varphi}^{-1}: \varphi(U \cap V) \times \mathbb{R}^n \to \psi(U \cap V) \times \mathbb{R}^n\) 
can be written explicitly using our knowledge of the differential for transition maps as
\(\tilde{\psi} \circ \tilde{\varphi}^{-1} (x^1, \dots, x^n, v^1, \dots, v^n) 
= \left( F^1(x), \dots, F^n(x), \frac{\partial F^1}{\partial x^j}(x) v^j, \dots, \frac{\partial F^n}{\partial x^j}(x) v^j \right).\),
where \( F = \tilde{\psi} \circ \tilde{\varphi}^{-1} \).
This is clearly smooth as the coordinate maps are and also

Choosing a countable cover \(\{U_i\}\) of \(M\) by smooth coordinate domains, 
we obtain a countable cover of \(TM\) by coordinate domains \(\{\pi^{-1}(U_i)\}\).
To check the Hausdorff condition (v), just note that any two points in the same fiber of \(\pi\) lie in one chart, 
while if \((p, v)\) and \((q, w)\) lie in different fibers, 
there exist disjoint smooth coordinate domains \(U, V\) for \(M\) such that \(p \in U\) and \(q \in V\), 
and then \(\pi^{-1}(U)\) and \(\pi^{-1}(V)\) are disjoint coordinate neighborhoods containing \((p, v)\) and \((q, w)\), respectively.

To see that \(\pi\) is smooth, note that with respect to charts \((U, \varphi)\) for \(M\) and \((\pi^{-1}(U), \tilde{\varphi})\) for \(TM\), 
its coordinate representation is \(\pi(x, v) = x\). The \( (x^1, \dots, x^n, v^1, \dots, v^n) \) are also referred to as natural coordinates.


Define the cotangent space

For any vector space \( V \) over a field \( \mathbb{K} \), we call linear functions \( f : V \to \mathbb{K} \)
covectors. The space under addition and multiplication is itself a vector space called the dual space \( V^\ast \).

We have in particular that for each finite dimensional space there is even an isomorphism \(\phi : V \to V^\ast \),
which is defined on the basis \( \{x_i\}_{i \in [\text{dim}\ V]} \) as \( \phi(x_i) = y_i \), s.t. \( y_i(x_j) = \delta_{i,j} \).

The cotangent space \( T^\ast_p\ M \) is then nothing else but the dual of the tangent space of a point.

We use the special notation \( dx^i \) for its basis vectors and similarly as for tangent vectors it holds for different
charts \( \phi, \psi \) inducing basis vectors \( d\tilde{x}^i \), \( d\tilde{y}^i \)

\( d\tilde{x}^k|_{\phi(p)} = \frac{dF^k}{d\tilde{x}^j}(\phi(p))dy^j|_{\phi(p)} \).


Define a section, local & global frames in the context of vector bundles

Let \(\pi: E \to M\) be a vector bundle. 

A local section of \(E\) is a continuous map \(\sigma: U \to E\) defined on some open subset \(U \subseteq M\) 
and satisfying \(\pi \circ \sigma = \text{Id}_U\). 
This means that \(\sigma(p)\) is an element of the fiber \(E_p\) for each \(p \in U\).
Should \( U = M \), we speak of a global section.
To emphasize the distinction, a section defined on all of \(M\) is sometimes called a global section.
We call \( \{ p \in M : \sigma(p) \neq 0 \} \) the support of a section.

If \(E \to M\) is a smooth vector bundle, 
the set of all smooth global sections of \(E\) is a vector space 
under pointwise addition and scalar multiplication:

\((c_1 \sigma_1 + c_2 \sigma_2)(p) = c_1 \sigma_1(p) + c_2 \sigma_2(p)\).

This vector space is usually denoted by \(\Gamma(E)\).
Further a section can act on a smooth function \( f \in C^{\infty}(M) \) by
\( (f\sigma)(p) = f(p)\sigma(p) \), which is again a new section (the original one multiplied with the value of \( f \) at \( p \)).
This will be useful in the context of vector fields.

\(p \in U\). 
A local frame for \(E\) over \(U\) is an ordered \(k\)-tuple \((\sigma_1, \dots, \sigma_k)\) of linearly independent local sections over \(U\),
that span \(E_p = \pi^{-1}(p)\) for each \( p \in U\); 
thus \((\sigma_1(p), \dots, \sigma_k(p))\) is a basis for the fiber \(E_p\), for each \(p \in U\). 
It is called a global frame if \(U = M\). If \(E \to M\) is a smooth vector bundle, 
a local or global frame is a smooth frame if each \(\sigma_i\) is a smooth section. 
We often denote a frame \((\sigma_1, \dots, \sigma_k)\) by \((\sigma_i)\).

Outline how local frames can generate local trivialization and vice versa

Local Frames Associated with Local Trivializations
Suppose \(\pi: E \to M\) is a smooth vector bundle. 
If \(\Phi: \pi^{-1}(U) \to U \times \mathbb{R}^k\) is a smooth local trivialization of \(E\), 
we can use the same idea as in the preceding example to construct a local frame for \(E\) over \(U\). 
Define maps \(\sigma_1, \dots, \sigma_k: U \to E\) by 

\(\sigma_i(p) = \Phi^{-1}(p, e_i) = \Phi^{-1} \circ \tilde{e}_i (p);\)

Then \(\sigma_i\) is smooth because \(\Phi\) is a diffeomorphism, 
and the fact that \(\pi_1 \circ \Phi = \pi\) implies that

\(\pi \circ \sigma_i(p) = \pi \circ \Phi^{-1}(p, e_i) = \pi_1(p, e_i) = p,\)

so \(\sigma_i\) is a section. 
To see that \((\sigma_i(p))\) forms a basis for \(E_p\), just note that \(\Phi\) restricts to an isomorphism from 
\(E_p\) to \(\{p\} \times \mathbb{R}^k\), and \(\Phi(\sigma_i(p)) = (p, e_i)\), so \(\Phi\) takes \((\sigma_i(p))\) 
to the standard basis for \(\{p\} \times \mathbb{R}^k \cong \mathbb{R}^k\). 
We say that this local frame \((\sigma_i)\) is associated with \(\Phi\).

Actually the converse holds as well that is
Every smooth local frame for a smooth vector bundle is associ-
ated with a smooth local trivialization

Proof. 
Suppose \(E \to M\) is a smooth vector bundle and \((\sigma_i)\) is a smooth local frame 
for \(E\) over an open subset \(U \subseteq M\). 
We define a map \(\Psi: U \times \mathbb{R}^k \to \pi^{-1}(U)\) by

\(\Psi(p, (v^1, \dots, v^k)) = v^i \sigma_i(p). \qquad (10.4)\)

The fact that \((\sigma_i(p))\) forms a basis for \(E_p\) at each \(p \in U\) implies that \(\Psi\) is bijective, 
and an easy computation shows that \(\sigma_i = \Psi \circ \tilde{e}_i\). 
Then with some effort we can show that \(\Psi\) is a diffeomorphism, so \(\Psi^{-1}\) is a smooth local trivialization, 
whose associated local frame is \((\sigma_i)\).

Combining these facts we have:
A smooth vector bundle is smoothly trivial if and only if it admits a smooth global frame.
(As a side effect this proves Lemma 4.4.4 from the lecture).


Define a vector field

With the setup of vector bundles a vector field is nothing else but a section 
of the tangent bundle, that is a smooth map \( \V : M \to TM \)
with \( \pi \circ V = \text{id}_{M} \).

In local coordinates a vector field can also be written as
\( V = v^i\frac{d}{dx^i} \)

We call the vector space of all vector fields \( \mathfrak{X}(M) \).
In particular vector fields have to surprising interactions with \( f \in C^\infty(M) \),
one is a little trivial namely creating a new vector field by just multiplying with the prefactor \( f \)
\( V : C^\infty(M) \times \mathfrak{X}(M) \to \mathfrak{X}(M), \ (f, V) \to f \cdot V \),
the other more interesting
\( V : C^\infty(M) \to C^\infty(M),\ (Vf)(p) \coloneqq D_{V(p)}f|_{p} \)
that is at each point \( p \), we take the directional derivative along the tangent vector
the vector field assigns to the point \( p \).
Making this more explicit we have
\( (Vf)(p) = (V^i_p \frac{d}{dx^i}|_{p})f = V^i_p \frac{df}{dx^i}|_{p}\)

So we observe that any vector field \( V \) can be considered as a derivation \( D_{V}|_{p} \) at \( p \),
We can generalize the pointwise notion of a derivation to say that \( D : C^\infty(M) \to C^\infty(M) \)
is a derivation, if its a derivation \( D(p) : C^\infty(M) \to \mathbb{R} \) at every point.

This is crucial, because it can be proven that such a map \( D \) is a derivation if and only if
\( Df = Xf \) for some smooth \( V \in \mathfrak{X}(M) \). This simplifies a lot of proofs from the lecture.

Define a Lie Group

A Lie group is a smooth manifold \( G \) (without boundary) that is also a group in the algebraic sense, 
with the property that the multiplication map \( m : G \times G \to G \) and inversion map \( i : G \to G \), given by
\(
m(g, h) = gh, \quad i(g) = g^{-1},
\)
are both smooth. 
A Lie group is, in particular, a topological group 
(a topological space with a group structure such that the multiplication and inversion maps are continuous).

If \(G\) is a Lie group, any element \(g \in G\) defines maps \(L_g, R_g: G \to G\), called left translation and right translation, respectively, by 

\(L_g(h) = gh, \quad R_g(h) = hg.\)

Because \(L_g\) can be expressed as the composition of smooth maps 

\(G \xrightarrow{t_g} G \times G \xrightarrow{m} G,\)

where \(t_g(h) = (g,h)\) and \(m\) is multiplication, it follows that \(L_g\) is smooth. 
It is actually a diffeomorphism of \(G\), because \(L_{g^{-1}}\) is a smooth inverse for it. 
Similarly, \(R_g: G \to G\) is a diffeomorphism. 
